The Talking Heads Experiment

### 2015-04-27 chap7/chap7.tex
final

Ludwig Wittgenstein went through two major 
periods in his thinking. During the first period he worked
within the framework of logical empiricism, initiated by 
Bertrand Russell and others at the end of the 19th century. 
The logicist approach defines a language by listing the 
elementary building blocks (the words), the combination 
rules (the syntax), and a mapping from words and 
syntactic structures to semantic interpretations (the semantics).  
It assumes that there is a universal, logical structure to the
world which can be captured once and for all in a
non-ambiguous logical language 
and that the permissible inferences can be catalogued 
exhaustively. This project has its roots in the 
17th century dream of
Leibnitz and Descartes to streamline rational thinking
so that it becomes as clear and non-controversial as
numerical calculation. 
Wittgenstein enthusiastically participated in this project, 
which is still vigorously being pursued today by logicians 
and linguists alike, even though many of them do not necessarily subscribe
to the universalist thesis motivating the early developments of 
logic.\footnote{Wittgenstein's main work from this period is \cite{Wittgenstein:1922}. 
A prototypical example of the logicist research program 
is found in \cite{Carnap:1928}.}

@@

What about synonymy? Let us look at the 
MF-diagram of [\textsc{g} 0.25–0.5], so that we can
see whether there are any other words in use for expressing
the same meaning (\figref{g02505.f}).
`fepi' has indeed emerged as dominant for this meaning, but 
this has not been without an intense struggle.
The tendency for synonymy to get damped is clearly present. 
Even though the lexicon occasionally destabilises, 
the lateral inhibition and the positive feedback loop between use 
and success causes self-organised MF-coherence. 
There is however something curious going on. 
In the early phases `xu' was dominant. Why did it 
destabilise and how has `fepi' has managed to become the
dominant expression for [\textsc{g} 0.25–0.5]?

>>>>

Ludwig Wittgenstein went through two major 
periods in his thinking. During the first period he worked
within the framework of logical empiricism, initiated by 
Bertrand Russell and others at the end of the 19th century. 
The logicist approach defines a language by listing the 
elementary building blocks (the words), the combination 
rules (the syntax), and a mapping from words and 
syntactic structures to semantic interpretations (the semantics).  
It assumes that there is a universal, logical structure to the
world which can be captured once and for all in a
non-ambiguous logical language 
and that the permissible inferences can be catalogued 
exhaustively. This project has its roots in the 
17th century dream of
Leibniz and Descartes to streamline rational thinking
so that it becomes as clear and non-controversial as
numerical calculation. 
Wittgenstein enthusiastically participated in this project, 
which is still vigorously being pursued today by logicians 
and linguists alike, even though many of them do not necessarily subscribe
to the universalist thesis motivating the early developments of 
logic.\footnote{Wittgenstein's main work from this period is \cite{Wittgenstein:1922}. 
A prototypical example of the logicist research program 
is found in \cite{Carnap:1928}.}

@@

What about synonymy? Let us look at the 
MF-diagram of [\textsc{g} 0.25–0.5], so that we can
see whether there are any other words in use for expressing
the same meaning (\figref{g0d2505.f}).
`fepi' has indeed emerged as dominant for this meaning, but 
this has not been without an intense struggle.
The tendency for synonymy to get damped is clearly present. 
Even though the lexicon occasionally destabilises, 
the lateral inhibition and the positive feedback loop between use 
and success causes self-organised MF-coherence. 
There is however something curious going on. 
In the early phases `xu' was dominant. Why did it 
destabilise and how has `fepi' has managed to become the
dominant expression for [\textsc{g} 0.25–0.5]?
